Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 PandaEvolution.tex
Iteration 4 and scattered editing

% -----------------------------------------------------------------------------
\subsection{Interfacing PanDA with Titan}
\label{ssec:panda-titan}

@@

Currently, very few HEP applications could benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows can
be offloaded from the Grid to Titan's large amount of cores. Further, when HEP
tasks can be partitioned to independent jobs, Titan worker nodes can be used to
execute up to 16 concurrent payloads, one for each available core. Given these
constraints and challenges, the type of task most suitable for execution at the
moment on Titan is Monte Carlo detector simulation. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and with
small input data requirements. Detector simulation tasks in ATLAS account for
$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
offloading.

@@

% -----------------------------------------------------------------------------
\subsection{PanDA Broker on Titan}
\label{ssec:panda_titan}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
    \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates
    the execution process of a detector simulation job, part of the production
    ATLAS MD workflow.}
  \end{center}
\label{fig:panda_broker}
\end{figure}

>>>>

% -----------------------------------------------------------------------------
%\subsection{Interfacing PanDA with Titan}
\subsection{Architectures and Interfaces}
\label{ssec:panda-titan}

@@

Currently, very few HEP applications could benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows can
be offloaded from the Grid to Titan's large amount of cores. Further, when HEP
tasks can be partitioned to independent jobs, Titan worker nodes can be used to
execute up to 16 concurrent payloads, one for each available core. Given these
constraints and challenges, the type of task most suitable for execution at the
moment on Titan is Monte Carlo detector simulation. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and with
small input data requirements. Detector simulation tasks in ATLAS are performed
via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
offloading.

@@

% -----------------------------------------------------------------------------
% \subsection{PanDA Broker on Titan}
\subsection{PanDA Broker}
\label{ssec:panda_titan}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
    \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates
    the execution process of a detector simulation job, part of the production
    ATLAS MD workflow.}
\label{fig:panda_broker}
\end{figure}

### 2017-04-03 overall.tex
Iteration 4 and scattered editing

We evaluate the efficiency, scalability and reliability of the PanDA Brokers
along two dimensions: (i) backfill utilization by the
PanDA Brokers; and (ii) the amount of runtime spent performing detector
simulations. We based these evaluations on measuring the number of cores
utilized by ATLAS on Titan, the overall backfill availability, the number of
detector simulations performed on all the resources available to ATLAS and on
Titan, and the failure rate of these simulations. All the measurements were
performed between January 2016 and February 2017, hereafter called `experiment
time window'.

@@

Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of core-hours
used by ATLAS on Titan during the experiment time window. ATLAS consumed a total
of 73.8M core-hours, for an average of $\approx$7M core-hours a month, with a
minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours in
February 2017.

@@

Figure~\ref{fig:backfill-utilization} (green bars) shows backfill utilization
during the experiment time window. Efficiency
(Figure~\ref{fig:backfill-utilization}, red labels) is defined as the fraction
of Titan’s cores available via backfill that was utilized by the PanDA Brokers.
The brokers reached 18\% average efficiency, with a minimum 8.9\% efficiency on
April 2016 and a maximum 33.5\% efficiency on February 2017. The number of total
backfill cores available was 38.1M in April 2016, and 33.1M in February 2017.
This shows that the measured difference in efficiency did not depend on a
comparable difference in total backfill availability.

@@

During the experiment time window, about 2.25M detector simulation jobs were
completed on Titan, for a total of 225M events processed. This is equivalent to
0.9\% of all the 250M detector simulations performed by ATLAS in the same period
of time, and 3.5\% of the 6.6B events processed by those jobs. Comparatively,
Titan contributed 3.9\% of the total of around 200K cores available to ATLAS on
Grid, Cloud, and HPC infrastructures together. These figures confirms the
relevance of supercomputers' resource contribution to the LHC Run 3, especially
when accounting for the amount of unused backfill availability and the rate of
improvement of PanDA efficiency.

@@

Figure~\ref{fig:backfill-distrib} shows the distribution of backfill
availability on Titan as a function of number of nodes and the time of their
availability (i.e., walltime). We recorded these data by polling Titan's Moab
scheduler at regular intervals during the experiment window time, while
developing and deploying PanDA Brokers. The mean number of nodes was 680, and
their mean walltime was 680 minutes. Detector simulations of 100 events, enable
to use down to 1 node for 1/2 of the mean walltime of backfill availability. As
such, it offers a good compromise for PanDA Broker efficiency.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/titan_backfill_avail.png}
    %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
    \caption{Backfill availability measures as walltime in minutes vs number of
    work nodes during the experiment window time. Availability was measured
    62175 times: mean number of work nodes available 680 (yellow line); mean
    walltime available 122 minutes (red line).}
\label{fig:backfill-distrib}
\end{figure}

Usually, detector simulations performed on the Grid process 10 times the number
of events than those on Titan. This explains the difference between the
percentage of detector simulations (9\%) and of events computed by those
simulations (3.9\%) on Titan. PanDA Broker could fit the number of events to the
walltime of the backfill availability on the base of the distribution of
Figure~\ref{fig:backfill-distrib}. That specific number of event could then be
pulled from the PanDA Event service~\cite{calafiura2015atlas} and given as input
to one or more simulations. Once packaged into the MPI script submitted to
titan's PBS batch system, these simulations would better fit backfill
availability, increasing the efficiency of PanDA Brokers.

@@

We use two main parameters to measure the performance of the detector simulation
jobs submitted to Titan: (i) the time taken to setup
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit~\cite{agostinelli2003geant4}; and (ii) the
distribution of the time taken by the Geant4 toolkit to simulate a certain
number of events.

AthenaMP has an initialization and configuration stage. At initialization time,
AthenaMP is assembled from a large number of shared libraries, depending on the
type of payload that will have to be computed. Once initialized, every algorithm
and service of AthenaMP is configured by a set of Python scripts. Both these
operations result in a large number of read operations on the filesystem,
including those required to access of small python scripts.

@@

\mtnote{TODO: specify that this is work in progress (see Sergey's comment) and
see how to move it to the data performance subsection}Initially, all the shared
libraries of AthenaMP and the python scripts for the configuration stage were
stored on the Spider 2 Lustre file system. However, the I/O patterns of the
initialization and configuration stages degraded the performance of the
filesystem (Figure~\ref{}\mtnote{We may want to aggregate some of the diagrams
about lustre's experiments here}). Since Spider 2 is a center-wide file system,
this resulted in  performance degradation for all OLCF resources and users.
\mtnote{I am afraid the details about the trace are too specific given the space
constraints of a SC submission. Please feel free to uncomment it if you
disagree.}

% As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
% exhibits a spike corresponding to the beginning of the runs before tapering
% off.

% \begin{minipage}{\linewidth}
% \begin{lstlisting}[language=bash,frame=single,basicstyle=\ttfamily\tiny,caption=ATLAS metadata trace,label=mdstrace]
% XK7 Application 9205593
%       39012 RPCs from 300 of 300 nodes
%         ~69291.96 per sec
%           37851 LDLM_ENQUEUE RPCs    ~67229.83 per sec
%                 pmin 13us pavg 42us pmax 4983us
%           611 LDLM_CANCEL RPCs    ~1085.24 per sec
%                 pmin 10us pavg 16us pmax 32us
%           277 MDS_CLOSE RPCs    ~492.00 per sec
%                 pmin 15us pavg 20us pmax 38us
%           154 MDS_READPAGE RPCs    ~273.53 per sec
%                 pmin 170us pavg 292us pmax 671us
%           86 MDS_GETXATTR RPCs    ~152.75 per sec
%                 pmin 15us pavg 21us pmax 65us
%           30 MDS_GETATTR RPCs    ~53.29 per sec
%                 pmin 16us pavg 20us pmax 27us
%           3 MDS_REINT RPCs    ~5.33 per sec
%                 pmin 103us pavg 136us pmax 196us
%           Overall times
%                 pmin 10us pavg 42us pmax 4983us
%
% XK7 Application 9205355
%       8698 RPCs from 62 of 62 nodes
%         ~15449.13 per sec
%           8445 LDLM_ENQUEUE RPCs    ~14999.76 per sec
%                 pmin 16us pavg 41us pmax 780us
%           92 MDS_CLOSE RPCs    ~163.41 per sec
%                 pmin 15us pavg 23us pmax 52us
%           55 MDS_READPAGE RPCs    ~97.69 per sec
%                 pmin 189us pavg 291us pmax 534us
%           52 LDLM_CANCEL RPCs    ~92.36 per sec
%                 pmin 12us pavg 16us pmax 25us
%           41 MDS_GETXATTR RPCs    ~72.82 per sec
%                 pmin 16us pavg 20us pmax 27us
%           13 MDS_GETATTR RPCs    ~23.09 per sec
%                 pmin 16us pavg 21us pmax 38us
%           Overall times
%                 pmin 12us pavg 42us pmax 780us
%
% \end{lstlisting}
% \end{minipage}

% As this problem was identified, the OLCF staff has enabled read-only access to
% certain NFS-exported directories from Titan compute nodes. This in turn
% allowed the OLCF staff to install a software package from a Titan login node
% and have it available read-only on a Titan compute node.

After careful characterization of the impact of ATLAS jobs on Lustre, the
performance degradation was addressed by moving the AthenaMP distribution to a
read-only NFS directory, shared among DTNs and worker nodes. This eliminated the
problem of metadata contention, improving metadata read performance from
$\approx$6,300 seconds on Lustre to $\approx$1,500 seconds on NFS. Further, at
configuration time, the input files of each detector simulation job were stored
to a ramdisk on the work nodes. This offered a marked improvement of reading
time: from 1,320 seconds on Lustre to 40 seconds on NFS.

%Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
%improvement on Titan. The circled region illustrates the switch from Lustre to
%NFS-exported directory for hosting the ATLAS release.

@@

@@

% \mtnote{Figure summarizing job performance on Titan analogous to those
% produced with NGE, as discussed in F2F meeting at Rutgers?}

% ----------------------------------------------------------------------------
% \subsection{Reliability of PanDA Broker and Detector Simulations on Titan}
% \label{ssec:reliability}

@@

% Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
% recorded by the PanDA Brokers between January 2016 and February 2017. The
% initial preponderance of unknown failures was progressively addressed by a
% more accurate failure model, specifically tailored to the PanDA Broker and
% Titan. Figure~\ref{fig:failures-titan} also shows the progressive improvement
% of the broker's reliability: In Februrary 2017, errors related to input files
% and initialization were almost eliminated and execution errors were comparable
% to those recorded in January 2017 even if almost twice as many simulations
% were performed (Figure~\ref{fig:backfill-utilization}). The increased amount
% of unknown failure in February indicates that furter work is needed to improve
% the PanDA Broker failure model.

% \begin{figure}[htp]
%     \includegraphics[clip,width=\columnwidth]{figures/failures_titan.pdf}
    % \caption{PanDA Broker failures during the experiment time window. Failures
    % were recorded with 52 exit codes, here aggregated into 6 categories. The
    % number of failures should be compared to the number of simulations
    % executed in each month as plotted in
    % Figure~\ref{fig:backfill-utilization}. Improvements in the failure model
    % of PanDA Broker shows a decrease in unknown failure and a the progressive
    % addressing of other categories of failure.}
% \label{fig:failures-titan}
% \end{figure}

% -----------------------------------------------------------------------------
\subsection{Characterizing PanDA I/O}

To better understand the I/O impact of ATLAS PanDA project on Titan
supercomputing environment we analyzed 1,175 jobs ran on the week of 10/25/2016,
for a total of 174 hours. Table~\ref{panda-olcf-stats} shows the overall
statistical breakdown of the observed file I/O impact of ATLAS at OLCF\@.
% Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written}
% show the file read and write I/O histograms for these 1,175 jobs,
% respectively. Figures~\ref{fig:atlas-titan-file-open}
% and~\ref{fig:atlas-titan-file-close} show the file $open()$ and $close()$
% metadata load histograms of the same  1,175 ATLAS jobs, respectively.
%
% As can be seen from Table~\ref{panda-olcf-stats},

ATLAS jobs use between 1 and 300 worker nodes, and 35 on average. 75\% of the
ATLAS jobs consume less than 25 nodes and 92\% less than 100. During the 174
hours of data collection, 6.75 ATLAS jobs were executed on average per hour,
each job running for an average of 1.74 hours.
% ATLAS jobs issues a large number of file read operations, as can be seen in
% Table~\ref{panda-olcf-stats}.
Every job read less than 250 GB and wrote less than 75 GB of data and, on
average, each job read 20 GB and wrote 6 GB of data.

\begin{table*}[t]
\centering
\begin{tabular}{lllllllll}
 & Num. Nodes & Duration (s) & Read (GB) & Written (GB) & GB Read/nodes & GB Written/nodes & $open()$ & $close()$ \\
Min & 1 & 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\
Max & 300 & 7,452 & 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\
Average & 35.66 & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 & 34,155.74 \\
Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376 & 231,346.55 & 53,799.08
\end{tabular}
\caption{The Statistical breakdown of the I/O impact of 1,175 PanDA jobs executed at OLCF for the week of 10/25/16}
\label{panda-olcf-stats}
\end{table*}

The I/O of the ATLAS job executed on Titan show an interesting pattern.
Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
worker node is less than 400 MB on average, while the amount of data written per
node is less than 170 MB on average. This correlates with our finding that ATLAS
PanDA jobs are read heavy. However,

ATLAS PanDA jobs are read heavy: On average, the amount of data read per worker
node is less than 400 MB, while the amount of data written is less than 170 MB.
This correlates with our finding that . However,
% as can be seen in Table~\ref{panda-olcf-stats} and figures
% {fig:atlas-titan-io-read} and {fig:atlas-titan-io-written},
the distributions of read and written data are quite different: The read
operation distribution per job shows a long tail, ranging from 12.5 GB to 250
GB, while the written amount of data has a very narrow distribution.\mtnote{Should we say why?}

The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
operations per second (not including file $stat()$ operations) and 5 file $close()$ operations per second, with similar distributions.
% As can be seen from figures~\ref{fig:atlas-titan-file-open}
% and~\ref{fig:atlas-titan-file-close}, they .
On average, the maximum number of file $open()$ operations per job is
$\approx$170/s and the maximum number of file $close()$ operations is
$\approx$39/s.
% on average per second per job.
For the 1,175 ATLAS PanDA jobs observed, the total number of file $open()$
operations is 172,089,760 and the total number of file $close()$ operations is
40,132,992. The difference between these two values is still under
investigation: One possible explanation is that ATLAS PanDA jobs don't call a
file $close()$ operation per every file $open()$ issued.

Overall, based on our experiments with ATLAS jobs, it can be safely concluded
that the file and metadata I/O load of ATLAS PanDA project on the OLCF Titan
supercomputing environment and the Spider 2 file system is not detrimental to
the center operations. At the current scale of the project, the overall impact
of ATLAS operations on Titan is minimal.

>>>>

We evaluate the efficiency, scalability and reliability of the deployment of PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and AthenaMP.
% two dimensions: (i) backfill utilization by the PanDA Brokers; and (ii) the
% amount of runtime spent performing detector simulations.
We based our characterizations on measuring the number of cores utilized by
ATLAS on Titan, the overall backfill availability, the number of detector
simulations performed on all the resources available to ATLAS and on Titan, and
the failure rate of these simulations. We also consider the the time taken to
setup AthenaMP and to simulate events in the ATLAS detector with it. All the
measurements were performed between January 2016 and February 2017, hereafter
called `experiment time window'.

\mtnote{TODO: expand the previous paragraph to include dimensions of AthenaMP characterization.}

@@

Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of Titan
core-hours used by ATLAS during the experiment time window. ATLAS consumed a
total of 73.8M Titan core-hours, for an average of $\approx$7M Titan core-hours
a month, with a minimum of 3.3M Titan core-hours in April 2016 and a maximum
14.8M Titan core-hours in February 2017.

@@

Figure~\ref{fig:backfill-utilization} (green bars) shows backfill utilization
during the experiment time window.  Efficiency
(Figure~\ref{fig:backfill-utilization}, red labels) is defined as the fraction
of Titan core-hours utilized by the PanDA Brokers (for ATLAS jobs) to the total
of Titan’s backfill availability. The brokers reached 18\% average efficiency,
with a minimum 8.9\% efficiency on April 2016 and a maximum 33.5\% efficiency on
February 2017. The total backfill available was 38.1M Titan core-hours in April
2016, and 33.1M Titan core-hours in February 2017. This shows that the measured
difference in efficiency did not depend on a comparable difference in total
backfill availability.

@@

During the experiment time window, about 2.25M detector simulation jobs were
completed on Titan, for a total of 225M events processed. This is equivalent to
0.9\% of all the 250M detector simulations performed by ATLAS in the same period
of time, and 3.5\% of the 6.6B events processed by those jobs. Comparatively,
Titan contributed 3.9\% of the total of around 200K cores available to ATLAS on
Grid, Cloud, and HPC infrastructures together. These figures confirms the
relevance of supercomputers' resource contribution to the LHC Run 2, especially
when accounting for the amount of unused backfill availability and the
improvement of PanDA efficiency across the experiment time window.

@@

Figure~\ref{fig:backfill-distrib} shows the distribution of backfill
availability on Titan as a function of number of nodes and the time of their
availability (i.e., walltime). We recorded these data by polling Titan's Moab
scheduler at regular intervals during the experiment window time, while
developing and deploying PanDA Brokers. The mean number of nodes was 691, and
their mean walltime was 126 minutes. Detector simulations of 100 events, enable
to use down to 1 node for 1/2 of the mean walltime of backfill availability. As
such, it offers a good compromise for PanDA Broker efficiency.

\begin{figure*}%[htp]
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_wnodes_distribution.pdf}
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_walltime_distribution.pdf}
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
    %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
    \caption{Backfill availability: distribution of number of work nodes (left); distribution of walltime in minutes (center); and their correlation (right). 62175 measures: mean number of work nodes available 691; mean
    walltime available 126 minutes.}
\label{fig:backfill-distrib}
\end{figure*}

Usually, detector simulations performed on the Grid process 10 times the number
of events than those on Titan. This explains the difference between the
percentage of detector simulations (9\%) and of events computed by those
simulations (3.9\%) on Titan. PanDA Broker could fit the number of events to the
walltime of the backfill availability on the base of the distributions of
Fig.~\ref{fig:backfill-distrib}. That specific number of event could then be
pulled from the PanDA Event service~\cite{calafiura2015atlas} and given as input
to one or more simulations. Once packaged into the MPI script submitted to
titan's PBS batch system, these simulations would better fit backfill
availability, increasing the efficiency of PanDA Brokers.

@@

We use two main parameters to measure the performance of the detector simulation
jobs submitted to Titan: (i) the time taken to setup AthenaMP; and (ii) the
distribution of the time taken by the Geant4 toolkit to simulate a certain
number of events.

AthenaMP has an initialization and configuration stage. At initialization time,
AthenaMP is assembled from a large number of shared libraries, depending on the
type of payload that will have to be computed. Once initialized, every algorithm
and service of AthenaMP is configured by a set of Python scripts. Both these
operations result in a large number of read operations on the filesystem shared
between the worker nodes and the DTNs, including the operations required to
access of small python scripts.

@@

% \mtnote{TODO: specify that this is work in progress (see Sergey's comment) and
% see how to move it to the data performance subsection}

Initially, all the shared libraries of AthenaMP and the python scripts for the
configuration stage were stored on the Spider 2 Lustre file system. However, the
I/O patterns of the initialization and configuration stages degraded the
performance of the filesystem.
% (Figure~\ref{}\mtnote{We may want to aggregate some of the diagrams about
% lustre's experiments here}).  Since Spider 2 is a center-wide file system,
% this resulted in  performance degradation for all OLCF resources and users.
% \mtnote{I am afraid the details about the trace are too specific given the
% space constraints of a SC  submission. Please feel free to uncomment it if
% you disagree.}
%
% As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
% exhibits a spike corresponding to the beginning of the runs before tapering
% off.
%
% \begin{minipage}{\linewidth}
% \begin{lstlisting}[language=bash,frame=single,basicstyle=\ttfamily\tiny,caption=ATLAS metadata trace,label=mdstrace]
% XK7 Application 9205593
%       39012 RPCs from 300 of 300 nodes
%         ~69291.96 per sec
%           37851 LDLM_ENQUEUE RPCs    ~67229.83 per sec
%                 pmin 13us pavg 42us pmax 4983us
%           611 LDLM_CANCEL RPCs    ~1085.24 per sec
%                 pmin 10us pavg 16us pmax 32us
%           277 MDS_CLOSE RPCs    ~492.00 per sec
%                 pmin 15us pavg 20us pmax 38us
%           154 MDS_READPAGE RPCs    ~273.53 per sec
%                 pmin 170us pavg 292us pmax 671us
%           86 MDS_GETXATTR RPCs    ~152.75 per sec
%                 pmin 15us pavg 21us pmax 65us
%           30 MDS_GETATTR RPCs    ~53.29 per sec
%                 pmin 16us pavg 20us pmax 27us
%           3 MDS_REINT RPCs    ~5.33 per sec
%                 pmin 103us pavg 136us pmax 196us
%           Overall times
%                 pmin 10us pavg 42us pmax 4983us
%
% XK7 Application 9205355
%       8698 RPCs from 62 of 62 nodes
%         ~15449.13 per sec
%           8445 LDLM_ENQUEUE RPCs    ~14999.76 per sec
%                 pmin 16us pavg 41us pmax 780us
%           92 MDS_CLOSE RPCs    ~163.41 per sec
%                 pmin 15us pavg 23us pmax 52us
%           55 MDS_READPAGE RPCs    ~97.69 per sec
%                 pmin 189us pavg 291us pmax 534us
%           52 LDLM_CANCEL RPCs    ~92.36 per sec
%                 pmin 12us pavg 16us pmax 25us
%           41 MDS_GETXATTR RPCs    ~72.82 per sec
%                 pmin 16us pavg 20us pmax 27us
%           13 MDS_GETATTR RPCs    ~23.09 per sec
%                 pmin 16us pavg 21us pmax 38us
%           Overall times
%                 pmin 12us pavg 42us pmax 780us
%
% \end{lstlisting}
% \end{minipage}
%
% As this problem was identified, the OLCF staff has enabled read-only access to
% certain NFS-exported directories from Titan compute nodes. This in turn
% allowed the OLCF staff to install a software package from a Titan login node
% and have it available read-only on a Titan compute node.
%
% After careful characterization of the impact of ATLAS jobs on Lustre, the
% performance degradation
This was addressed by moving the AthenaMP distribution to a read-only NFS
directory, shared among DTNs and worker nodes. NFS eliminated the problem of
metadata contention, improving metadata read performance from $\approx$6,300
seconds on Lustre to $\approx$1,500 seconds on NFS.

% Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
% improvement on Titan. The circled region illustrates the switch from Lustre to
% NFS-exported directory for hosting the ATLAS release.

@@

Once set up, AthenaMP is used to execute 16 concurrent Geant4 simulators on each
of Titan's worker node. Geant4 requires to read events descriptions from a
filesystem and simulate them as they would happen within the ATLAS detector. We
characterized both the compute performance of the simulation and the impact of
acquiring event descriptions on the filesystem.

@@

% \mtnote{Figure summarizing job performance on Titan analogous to those
% produced with NGE, as discussed in F2F meeting at Rutgers?}
%
% ----------------------------------------------------------------------------
% \subsection{Reliability of PanDA Broker and Detector Simulations on Titan}
% \label{ssec:reliability}
%
% Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
% recorded by the PanDA Brokers between January 2016 and February 2017. The
% initial preponderance of unknown failures was progressively addressed by a
% more accurate failure model, specifically tailored to the PanDA Broker and
% Titan. Figure~\ref{fig:failures-titan} also shows the progressive improvement
% of the broker's reliability: In Februrary 2017, errors related to input files
% and initialization were almost eliminated and execution errors were comparable
% to those recorded in January 2017 even if almost twice as many simulations
% were performed (Figure~\ref{fig:backfill-utilization}). The increased amount
% of unknown failure in February indicates that furter work is needed to improve
% the PanDA Broker failure model.
%
% \begin{figure}[htp]
%     \includegraphics[clip,width=\columnwidth]{figures/failures_titan.pdf}
    % \caption{PanDA Broker failures during the experiment time window. Failures
    % were recorded with 52 exit codes, here aggregated into 6 categories. The
    % number of failures should be compared to the number of simulations
    % executed in each month as plotted in
    % Figure~\ref{fig:backfill-utilization}. Improvements in the failure model
    % of PanDA Broker shows a decrease in unknown failure and a the progressive
    % addressing of other categories of failure.}
% \label{fig:failures-titan}
% \end{figure}
%
% -----------------------------------------------------------------------------
% \subsection{Characterizing PanDA I/O}
%
% To better understand the I/O impact of ATLAS PanDA project on Titan
% supercomputing environment

We studied the impact of acquiring event descriptions on Lustre by analyzing
1,175 jobs ran on the week of 10/25/2016, for a total of 174 hours.
Table~\ref{panda-olcf-stats} shows the overall statistical breakdown of the
observed file I/O.
% Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written}
% show the file read and write I/O histograms for these 1,175 jobs,
% respectively. Figures~\ref{fig:atlas-titan-file-open}
% and~\ref{fig:atlas-titan-file-close} show the file $open()$ and $close()$
% metadata load histograms of the same  1,175 ATLAS jobs, respectively.
%
% As can be seen from Table~\ref{panda-olcf-stats},
%
ATLAS used between 1 and 300 worker nodes, and 35 on average. 75\% of the jobs
run by ATLAS consumed less than 25 nodes and 92\% less than 100. During the 174
hours of data collection, 6.75 ATLAS jobs were executed on average per hour,
each job running for an average of 1.74 hours.
% ATLAS jobs issues a large number of file read operations, as can be seen in
% Table~\ref{panda-olcf-stats}.
Every job read less than 250 GB and wrote less than 75 GB of data and, on
average, each job read 20 GB and wrote 6 GB of data.

\begin{table*}[t]
\centering
\begin{tabular}{lllllllll}
 & Num. Nodes & Duration (s) & Read (GB) & Written (GB) & GB Read/nodes & GB Written/nodes & $open()$ & $close()$ \\
Min & 1 & 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\
Max & 300 & 7,452 & 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\
Average & 35.66 & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 & 34,155.74 \\
Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376 & 231,346.55 & 53,799.08
\end{tabular}
\caption{The Statistical breakdown of the I/O impact of 1,175 jobs ATLAS executed at OLCF for the week of 10/25/16}
\label{panda-olcf-stats}
\end{table*}

% The I/O of the ATLAS jobs show an interesting pattern.
% Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
% worker node is less than 400 MB on average, while the amount of data written
% per node is less than 170 MB on average. This correlates with our finding
% that ATLAS PanDA jobs are read heavy. However,

ATLAS jobs are read heavy: On average, the amount of data read per worker node
is less than 400 MB, while the amount of data written is less than 170 MB.
% as can be seen in Table~\ref{panda-olcf-stats} and figures
% {fig:atlas-titan-io-read} and {fig:atlas-titan-io-written},
Distributions of read and written data are different: The read operation
distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
the written amount of data has a very narrow distribution.

The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
operations per second (not including file $stat()$ operations) and 5 file $close()$ operations per second, with similar distributions.
% As can be seen from figures~\ref{fig:atlas-titan-file-open}
% and~\ref{fig:atlas-titan-file-close}, they .
On average, the maximum number of file $open()$ operations per job is
$\approx$170/s and the maximum number of file $close()$ operations is
$\approx$39/s.
% on average per second per job.
For the 1,175 ATLAS jobs observed, the total number of file $open()$ operations
is 172,089,760 and the total number of file $close()$ operations is 40,132,992.
The difference between these two values is still under investigation: One
possible explanation is that ATLAS jobs don't call a file $close()$ operation
per every file $open()$ issued.

% based on our experiments with ATLAS jobs, it can be safely concluded
% that the file and metadata I/O load of ATLAS PanDA project on the OLCF Titan
% supercomputing environment and the Spider 2 file system is not detrimental to
% the center operations. At the current scale of the project, the overall impact
% of ATLAS operations on Titan is minimal.

Overall, the average time taken to read events from input files stored on Lustre is 1,320, comparable to the time taken to read the file required by assembling AthenaMP from NFS. Preliminary investigation shows that this time could be reduced to 40 seconds by loading the event descriptions into the RAM disk available on each worker node. Events descriptions could be transferred from Lustre to the RAM disk while configuring and initializing AthenaMP, almost halving the time currently required by initiating a Genat4 simulation.

@@
